---
title: "Class 14 DESeq2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load in the data

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

## how many genes do we have?
```{r}
nrow(counts)
```

Compare control vs treated count data contained in 'counts' for each gene (rows)

using $ to get the dex column which stores control or treated groups, == to check to see if it's TRUE, and then access the id's of the controls. Store this in control value

```{r}
metadata$dex == "control"

control <- metadata[metadata$dex == "control", ]
control
```
```{r}
counts[ ,control$id]
```

Now find the mean value for each gene (ie each row)

```{r}
control.mean <- rowSums(counts[ ,control$id])/(length(control$id))
names(control.mean) <- counts$ensgene
```

# Do the same thing except for the treated samples

```{r}
treated <- metadata[metadata$dex == "treated", ]
treated
counts[ ,treated$id]
#could also just use rowmeans here
treated.mean <- rowSums(counts[ ,treated$id])/(length(treated$id))
names(treated.mean) <- counts$ensgene
```

```{r}
mycounts <- data.frame(control.mean, treated.mean)
```

```{r}
plot(mycounts)
```

```{r}
hist(mycounts$control.mean, breaks = 300)
```



```{r}
plot(log10(mycounts))
```

# Removing genes with zero counts, practicing with toy dataset "x"

```{r}
x <- data.frame(c(1,3,10,0),
                c(1,3,0,0))
x==0

unique(which(x==0, arr.ind = TRUE)[,"row"])

```
 #using the above code on mycounts
 
```{r}
to_rm <- unique(which(mycounts==0, arr.ind = TRUE)[,"row"])
newcounts <- mycounts[-to_rm,]
nrow(newcounts)
```
 
Calculating log2FC, and adding it to meancounts data.frame
```{r}
newcounts$log2fc <- log2(newcounts[,"treated.mean"]/newcounts[,"control.mean"])
head(newcounts)
```

#how many genes have a log2fc of 2?

```{r}
sum(newcounts$log2fc > 2)
```
#how many genes have a log2fc of -2?

```{r}
sum(newcounts$log2fc < -2)
```

Let's use DESeq2

```{r}
library(DESeq2)
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds

```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

## Volcano Plots

plot of lg2fc residues

```{r}
plot(res$log2FoldChange, -log10(res$padj), col = "gray")
abline(v=c(-2,+2), lty =2)
abline(h=-log(0.05), lty = 2)
```


