---
title: "Class10"
author: "Joseph Herdy"
date: "2/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Class 10, it's like PCA on cancer cells or something today...?

loading the wisconsin cancer biopsy data I downloaded from the lecture 10 worksheet

```{r}
wisc_df <- read.csv(file = "WisconsinCancer.csv")
head(wisc_df)
```

```{r}
wisc_data <- as.matrix(wisc_df[,3:32])
head(wisc_data)
```

> Q how many patients are in the dataset? Need the number of rows

> Q how many cancer and noncancer patients are there?

```{r}
nrow(wisc_data)
```

```{r}
table(wisc_df$diagnosis)
```

> Q how many column names have "mean"?

take length of grep of our mean pattern to get the number of mean containing columns
```{r}
length(grep("mean", colnames(wisc_data)))
```

## enter PCA

gotta check to see if the data needs to be scaled first.

us apply to run function sd on all the columns (2) in wisc_data
rounding the result so it's easier to look at ()
```{r}
round(apply(wisc_data,2,sd))
```


# Perform PCA on wisc_data 
```{r}
wisc.pr <- prcomp(wisc_data, scale = TRUE)
summary(wisc.pr)
```

plotting out this PCA data. biplot sucks big dicks apparently

```{r}
attributes(wisc.pr)
```

the pc's are stored in the component "x"

```{r}
plot(wisc.pr$x[,1:2], col=wisc_df$diagnosis)
```

Looks like PC1 is separating malignant from benign ayyyyy boiiiiiiiiiiii

# Calculate variance of each component

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```



Calculate the variance explained by each principal component by dividing by the total variance explained of all principal components. Assign this to a variable called pve and create a plot of variance explained for each principal component.

# Variance explained by each principal component: pve

```{r}
pve <- pr.var / sum(pr.var)
pve 

```

# Plot variance explained for each principal component
```{r}
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```


# Alternative scree plot of the same data, note data driven y-axis
```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```



```{r}
wisc.pr.hc <- hclust(dist(wisc.pr$x[,1:3]), method = "ward.D2")
plot(wisc.pr.hc)
```














```{r}
grps <- cutree(wisc.pr.hc, h = 80)
```


```{r}
table(grps, wisc_df$diagnosis)
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = grps)
```


## We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.
```{r}
url <- "https://tinyurl.com/new-samples-CSV"

new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```




```{r}
plot(wisc.pr$x[,1:2], col=wisc_df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2],labels = c(1,2), col="white" )
```











